### 2025-04-10 correlation.qmd
20250410 - placeholder
20250410 - update figures

```{r}
#| label: fig-anscombeQuartet
#| fig-cap: "Anscombe's Quartet."
#| fig-alt: "Anscombe's Quartet."
#| code-fold: true

@@

Consider the following association between air yards share and receiving yards:

```{r}
cor.test(
  player_stats_seasonal$air_yards_share,
  player_stats_seasonal$receiving_yards
)
```

There is a statistically significant, moderate positive association between air yards share and receiving yards, as depicted in @fig-scatterplotWithoutRangeRestriction.

```{r}
#| label: fig-scatterplotWithoutRangeRestriction
#| fig-cap: "Scatterplot of Player Receiving Yards (Season) by Air Yard Share. The best-fit line is in black."
#| fig-alt: "Scatterplot of Player Receiving Yards (Season) by Air Yard Share. The best-fit line is in black."
#| code-fold: true

plot_scatterplotWithoutRangeRestriction <- ggplot2::ggplot(
  data = player_stats_seasonal %>% filter(air_yards_share > 0),
  aes(
    x = air_yards_share,
    y = receiving_yards)) +
  geom_point(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season), # add season for mouse over tooltip
    alpha = 0.10) +
  geom_smooth(
    method = "lm",
    color = "black") +
  #geom_smooth() +
  coord_cartesian(
    ylim = c(0,NA),
    expand = FALSE) +
  labs(
    x = "Air Yards Share",
    y = "Receiving Yards",
    title = "Receiving Yards (Season) by Air Yards Share",
    subtitle = "(Among Players with Air Yards Share > 0)"
  ) +
  theme_classic()

@@

Now, consider the same association when we restrict the range to examine only those players whose air yards share was more than 25 (thus resulting in less variability):

```{r}
cor.test(
  player_stats_seasonal %>% 
    filter(air_yards_share > 25) %>% 
    select(air_yards_share) %>% 
    pull(),
  player_stats_seasonal %>% 
    filter(air_yards_share > 25) %>% 
    select(receiving_yards) %>% 
    pull()
)
```

The association is no longer statistically significant, and the association is weak in terms of [effect size](#sec-effectSizeThresholds), as depicted in @fig-scatterplotWithRangeRestriction.

```{r}
#| label: fig-scatterplotWithRangeRestriction
#| fig-cap: "Scatterplot of Player Receiving Yards (Season) by Target Share. The linear best-fit line is in black. The nonlinear best-fit line is in blue."
#| fig-alt: "Scatterplot of Player Receiving Yards (Season) by Target Share. The linear best-fit line is in black. The nonlinear best-fit line is in blue."
#| code-fold: true

plot_scatterplotWithRangeRestriction <- ggplot2::ggplot(
  data = player_stats_seasonal %>% filter(air_yards_share > 25),
  aes(
    x = air_yards_share,
    y = receiving_yards)) +
  geom_point(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season)) + # add season for mouse over tooltip)
  geom_smooth(
    method = "lm",
    color = "black") +
  #geom_smooth() +
  coord_cartesian(
    ylim = c(0,NA),
    expand = FALSE) +
  labs(
    x = "Air Yards Share",
    y = "Receiving Yards",
    title = "Receiving Yards (Season) by Air Yards Share",
    subtitle = "(Among Players with Air Yards Share > 25)"
  ) +
  theme_classic()

>>>>

```{r}
#| label: fig-anscombeQuartet
#| fig-cap: "Anscombe's Quartet."
#| fig-alt: "Anscombe's Quartet."
#| fig-height: 6
#| fig-width: 6
#| code-fold: true

@@

Consider the following association between passing attempts and expected points added (EPA) via passing.
Expected points added from a given play is calculated as the difference between a team's expected points before the play and the team's expected points after the play.
This can be summed across all passing plays to determine a player's EPA from passing during a given season.

```{r}
cor.test(
  player_stats_seasonal$attempts,
  player_stats_seasonal$passing_epa
)
```

There is a statistically significant, moderate positive association between passing attempts and expected points added (EPA) via passing, as depicted in @fig-scatterplotWithoutRangeRestriction.

```{r}
#| label: fig-scatterplotWithoutRangeRestriction
#| fig-cap: "Scatterplot of Player Passing Expected Points Added (Season) by Passing Attempts, Without Range Restriction. The best-fit line is in blue."
#| fig-alt: "Scatterplot of Player Passing Expected Points Added (Season) by Passing Attempts, Without Range Restriction. The best-fit line is in blue."
#| code-fold: true

plot_scatterplotWithoutRangeRestriction <- ggplot2::ggplot(
  data = player_stats_seasonal,
  aes(
    x = attempts,
    y = passing_epa)) +
  geom_point(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season), # add season for mouse over tooltip
    alpha = 0.3) +
  geom_smooth(
    method = "lm") +
  #geom_smooth() +
  coord_cartesian(
    expand = FALSE) +
  labs(
    x = "Player's Passing Attempts",
    y = "Player's Expected Points Added (EPA) via Passing",
    title = "Player's Passing Expected Points Added (EPA)\nby Passing Attempts (Season)",
    #subtitle = ""
  ) +
  theme_classic()

@@

Now, consider the same association when we restrict the range to examine only those players who had fewer than 450 pass attempts in a season (thus resulting in less variability):

```{r}
cor.test(
  player_stats_seasonal %>% 
    filter(attempts < 450) %>% 
    select(attempts) %>% 
    pull(),
  player_stats_seasonal %>% 
    filter(attempts < 450) %>% 
    select(passing_epa) %>% 
    pull()
)
```

The association is no longer positive and is weak in terms of [effect size](#sec-effectSizeThresholds), as depicted in @fig-scatterplotWithRangeRestriction.

```{r}
#| label: fig-scatterplotWithRangeRestriction
#| fig-cap: "Scatterplot of Player Passing Expected Points Added (Season) by Passing Attempts, With Range Restriction. The best-fit line is in blue."
#| fig-alt: "Scatterplot of Player Passing Expected Points Added (Season) by Passing Attempts, With Range Restriction. The best-fit line is in blue."
#| code-fold: true

plot_scatterplotWithRangeRestriction <- ggplot2::ggplot(
  data = player_stats_seasonal %>% filter(attempts < 450),
  aes(
    x = attempts,
    y = passing_epa)) +
  geom_point(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season), # add season for mouse over tooltip
    alpha = 0.3) +
  geom_smooth(
    method = "lm") +
  #geom_smooth() +
  coord_cartesian(
    expand = FALSE) +
  labs(
    x = "Player's Passing Attempts",
    y = "Player's Expected Points Added (EPA) via Passing",
    title = "Player's Passing Expected Points Added (EPA)\nby Passing Attempts (Season)",
    subtitle = "(Among Players with < 450 Passing Attempts)"
  ) +
  theme_classic()
